---
title: "Pivot Lobsters"
author: "Alanna Ramirez"
date: "1/27/2020"
output: html_document
---

Attaching our packages and reading in data
```{r}
library(tidyverse) #install.packages("tidyverse")
library(readxl)
library(here)
 # install.packages("skimr")
library(skimr)
library(kableExtra) # install.packages("kableExtra")
```
# read in lobster data

lobsters <- read_xlsx(here("data", "lobsters.xlsx"),
skip = 4)
```{r}
lobsters <- read_xlsx(here("data", "lobsters.xlsx"),
                      skip = 4)
```

Viewing with our skimr

```{r}
# these are equivalent
skimr::skim(lobsters)
skim(lobsters) # not explicit

```

# Pivot Tables
In R, Pivot tables are: `group_by() %>% summarize()`
`n()` function will count observations

```{r}
# group by year
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())

# group by year and site
lobsters %>%
  group_by(year, site) %>%
  summarize(count_by_year_site = n())

# group by site and year
lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_site_year = n())
```

Summarize by multiple variables
And assign as an object
```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

Table formatting with `kableExtra::kable()`
```{r}
siteyear_summary %>%
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.

## Activity
1. Build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm` and a function called `median`)
2. create a ggplot and save. For example:

3. sync this to github

```{r}
median_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            median_size_mm = median(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
ggplot(data = median_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))

```

